---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simrabid

`simrabid` is an implementation of a spatially explicit, individual-based model of canine rabies. It is very much a work in progress!

## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mrajeev08/simrabid")

# To install suggested packages & dependencies:
devtools::install_github("mrajeev08/simrabid", dependencies = TRUE)
```
## Minimum Working Example

This is a basic example which shows you the very basics:

```{r example}
library(raster)
library(data.table)
library(sf)
library(tidyr)
library(dplyr)
library(magrittr)
library(simrabid)
library(ggplot2)

# Use the example serengeti shapefile in the package
sd_shapefile <- st_read(system.file("extdata/sd_shapefile.shp", 
                                    package = "simrabid"))

# 1. set up the space at 1000m res
sd_shapefile$id_col <- 1:nrow(sd_shapefile)
out <- setup_space(shapefile = sd_shapefile, resolution = 1000, id_col = "id_col", 
                   use_fasterize = TRUE)

# Fake some pop data here
pop_out <- out
values(pop_out) <- rpois(ncell(pop_out), 20)
pop_out[is.na(out)] <- NA
plot(pop_out)

# 2. set-up simulation framework (timesteps + demography)
start_up <- setup_sim(start_date = "2002-01-01",
                      apprx_end_date = "2020-01-01", 
                      days_in_step = 7, 
                      rast = out, 
                      death_rate_annual = 0.48, 
                      birth_rate_annual = 0.52,
                      waning_rate_annual = 1/3,
                      params = list(start_pop = pop_out[]), 
                      by_admin = FALSE)

# Set up vaccination
vacc_dt <- simrabid::sim_campaigns(locs = 1:75, campaign_prob = 0.9, 
                                   coverage = 0.7, sim_years = 20, 
                                   burn_in_years = 5,
                                   steps_in_year = 52)
vacc_dt_none <- vacc_dt[0]

# Without vax
system.time({
  test <- simrabid(start_up, start_vacc = 0.2, I_seeds = 3,
                 vacc_dt = vacc_dt_none,
                 params = c(list(R0 = 1.1, k = 1, iota = 0.25),
                            param_defaults),
                 days_in_step = 7,
                 observe_fun = beta_detect_monthly,
                 serial_fun = serial_lognorm,
                 dispersal_fun = dispersal_lognorm,
                 secondary_fun = nbinom_constrained,
                 incursion_fun = sim_incursions_pois,
                 movement_fun = sim_movement_continuous,
                 sequential = FALSE, allow_invalid = TRUE,
                 leave_bounds = TRUE, max_tries = 100,
                 summary_fun = return_env,
                 track = FALSE,
                 weights = NULL,
                 row_probs = NULL,
                 coverage = TRUE,
                 break_threshold = 0.8, 
                 by_admin = FALSE) 
}
)

ggplot(test$I_dt) + geom_bar(aes(x = floor(t_infectious)))

formals(use_mget)$names <- c("I_dt", "N_mat")
system.time({
  test <- simrabid(start_up, start_vacc = 0.2, I_seeds = 3,
                 vacc_dt = vacc_dt,
                 params = c(list(R0 = 1.2, k = 1, iota = 1),
                            param_defaults),
                 days_in_step = 7,
                 observe_fun = beta_detect_monthly,
                 serial_fun = serial_lognorm,
                 dispersal_fun = steps_weibull,
                 secondary_fun = nbinom_constrained,
                 incursion_fun = sim_incursions_pois,
                 movement_fun = sim_movement_continuous,
                 sequential = TRUE, allow_invalid = TRUE,
                 leave_bounds = TRUE, max_tries = 100,
                 summary_fun = use_mget,
                 track = TRUE,
                 weights = NULL,
                 row_probs = NULL,
                 coverage = TRUE,
                 break_threshold = 0.8, 
                 by_admin = FALSE) 
}
)

# With vax
ggplot(test$I_dt) + geom_bar(aes(x = floor(t_infectious)))
plot(colSums(test$N_mat))
```

`simrabid` is written to be modular, feel free to poke around the function documentation to see the options--technical vignettes coming

## Roadmap

- A relative probability based movement model (i.e. movement probability is relative 
  to index location, for instance if you want to account for landscape barriers (roads, rivers, etc.)); 
  see old commit code for how this might work with a probability list;

- Use non utm coordinate space, instead use cellFromXY in raster and haversine distances 
 to get location of cell id movements so that you can simulate across larger 
 spatial scales; one issue is then your scale of aggregation gets distorted, aggregate grid cells somehow
 so it's approximately 1 km?

- Implement carrying capacity on pop growth, and replacement of dogs removed due to infection.

- Implement colonization of uninhabited patches (with some limits so households can't pop up in space
  that is uninhabitable, i.e. rivers/roads/etc.)

- Build a constructor class that gets passed to simrabid function so that only **valid** combinations
  of model arguments can be passed, and you only have to test this once for N simulations. 

- Profile and speed up! 
  - Vaccination function
  - Filtering data.table
  - Only track currently infectious + exposed linelist
  - Easy fixes (i.e. storing things in the appropriate type, keys onf filters, etc.)

- Issue with intermediate scales where small admin units do not get matched to 
  any grid cell

- Village metapopulation (separate function?)

- Example output environment for customizing summary functions

- Applying mortality to exposed class?

- Documentation on how to use & customize

- Construct synthetic populations or use high res pop data to get estimates of spatial dog pops (popcompr?)

- Benchmarks across scales, etc.

## Bug fixes

- Issue with setup_space, id col has to be numeric

## Known limitations / potential extensions
- Doesn't simulate expansion of occupied cells (i.e. colonization of patches by doggos)
- Carrying capacity for growth
